### 2016-05-26 writing/latex/introduction.tex
Make changes suggested by Michael to thesis. Bjarki's pedantry in presentation.

\section{Mesh Patterns}
When looking at a classical pattern, for example \(\perm{2,3,1}\), any occurrences can be arbitrarily
placed. However it may be interesting to consider occurrences where the elements
corresponding to the \(2\) and \(3\) are adjacent in the permutation. In the past
these sort of questions have lead to a variety of definitions.  \textcite{babstein2000}
considered \emph{vincular} patterns (also known as \emph{generalised} or \emph{dashed} patterns),
those where two adjacent entries in the pattern must be adjacent in the permutation,
\ie \(R\) is a union of vertical strips.
\textcite{MR2652101} look at classes of pattern where both columns and rows
can be shaded, these are called \emph{bivincular} patterns.
\emph{Bruhat-restricted} patterns were studied by \textcite{MR2264071} in order
to establish necessary conditions for a Schubert variety to be Gorenstein.
Mesh patterns also encompass a subset of \emph{barred} patterns introduced by
\textcite{MR2716312}, those with only one barred letter.

@@

The example of an occurrence of \(\perm{2,3,1}\) with adjacent elements \(31\)
can be represented as a mesh pattern by
\begin{equation*}
  \pattern{}{2,3,1}{1/0,1/1,1/2,1/3}
\end{equation*}

@@

Two patterns are said to be \emph{coincident} if they avoid the same set of
permutations and \emph{Wilf-equivalent} if they avoid the same number of permutations
at every length.

>>>>

\section{Mesh Patterns}
When looking at a classical pattern, for example \(\perm{2,3,1}\), any occurrences can be arbitrarily
placed. However it may be interesting to consider occurrences where the elements
corresponding to the \(2\) and \(3\) are adjacent in the permutation. In the past
these sort of questions have lead to a variety of definitions.  \textcite{babstein2000}
considered \emph{vincular} patterns (also known as \emph{generalised} or \emph{dashed} patterns),
those where two adjacent entries in the pattern must be adjacent in the permutation.
\textcite{MR2652101} look at classes of pattern where both columns and rows
can be shaded, these are called \emph{bivincular} patterns.
\emph{Bruhat-restricted} patterns were studied by \textcite{MR2264071} in order
to establish necessary conditions for a Schubert variety to be Gorenstein.
Mesh patterns also encompass a subset of \emph{barred} patterns introduced by
\textcite{MR2716312}, those with only one barred letter.

@@

The example of an occurrence of \(\perm{2,3,1}\) with adjacent elements \(23\)
can be represented as a mesh pattern by
\begin{equation*}
  \pattern{}{2,3,1}{1/0,1/1,1/2,1/3}
\end{equation*}

@@

Two patterns are said to be \emph{coincident} if they are avoided by the same set of
permutations and \emph{Wilf-equivalent} if they are avoided by the same number of permutations
at every length.

### 2016-05-26 writing/latex/wilf.tex
Make changes suggested by Michael to thesis. Bjarki's pedantry in presentation.

Computing avoiders up to length \(10\) suggest that there are at least \(23\)
Wilf-classes, of which \(13\) are non-trivial.

@@

\subsection{}
\nextvar
The following patterns are experimentally Wilf-equivalent up to length 10 in
\(\av{\perm{2,3,1}}\)
\begin{equation*}
    \pattern{}{1,2}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
    \pattern{}{1,2}{1/1,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,1/1,2/1,2/2},
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,1/1,2/2},
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,1/1},
    \pattern{}{2,1}{0/0,0/1,0/2,1/1,1/2,2/0,2/1,2/2}
\end{equation*}
Consider containers of these patterns.
For each of these patterns there is precisely one occurrence
in any permutation containing the pattern. Now consider
the points in the region corresponding to the unshaded box in each case. Each
must contain an avoider of \(\perm{2,3,1}\) that is of length
\(n-2\). Therefore these classes are all Wilf-equivalent
and the number of length \(n\) avoiders is
\begin{equation*}
    \var_n = C_n - C_{n-2}
\end{equation*}
for \(n \ge 2\) where \(C_n\) is the \(n\)th Catalan number, the number of
\(\perm{2,3,1}\) avoiders of length \(n\). This gives the sequence
\begin{equation*}
    1, 1, 1, 4, 12, 37, 118, 387, 1298, 4433, 15366,\dotsc \tag{\(C_n - \)\href{https://oeis.org/A001453}{A001453} offset 2}
\end{equation*}

>>>>

Computing avoiders up to length \(10\) suggest that there are at least \(23\)
Wilf-classes, of which \(13\) are non-trivial, this means that there are
Wilf-equivalences between patterns that are not explained by coincidences.

@@

\subsection{}
\nextvar
The following patterns are experimentally Wilf-equivalent up to length 10 in
\(\av{\perm{2,3,1}}\)
\begin{equation*}
    \pattern{}{1,2}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
    \pattern{}{1,2}{1/1,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,1/1,2/1,2/2},
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,1/1,2/2},
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,1/1},
    \pattern{}{2,1}{0/0,0/1,0/2,1/1,1/2,2/0,2/1,2/2}
\end{equation*}
Consider containers of these patterns in \(\av{\perm{2,3,1}}\).
For each of these patterns there is precisely one occurrence
in any permutation containing the pattern. Now consider
the points in the region corresponding to the unshaded box in each case. Each
must contain an avoider of \(\perm{2,3,1}\) that is of length
\(n-2\). Therefore these classes are all Wilf-equivalent
and the number of length \(n\) avoiders is
\begin{equation*}
    \var_n = C_n - C_{n-2}
\end{equation*}
for \(n \ge 2\) where \(C_n\) is the \(n\)th Catalan number, the number of
\(\perm{2,3,1}\) avoiders of length \(n\). This gives the sequence
\begin{equation*}
    1, 1, 1, 4, 12, 37, 118, 387, 1298, 4433, 15366,\dotsc \tag{\(C_n - \)\href{https://oeis.org/A001453}{A001453} offset 2}
\end{equation*}

### 2016-05-26 writing/presentation/coincidences.tex
Make changes suggested by Michael to thesis. Bjarki's pedantry in presentation.

\begin{frame}{Second Dominating rule}
  \begin{block}{}
    The patterns
    \begin{equation*}
    m_1 = \pattern{}{2,1}{0/0,0/2,1/0,2/1} \text{ and } m_2 = \pattern{}{2,1}{0/0,0/2,1/0,1/1,2/1}
    \end{equation*}
    are coincident in \(\av{\perm{2,3,1}}\).
  \end{block}
\end{frame}

@@

\begin{frame}
  \begin{example}
  Considering \(m_1\) again
  \begin{equation*}
    \begin{tikzpicture}[scale=\picscale]
        {\only<3-4> \fill[pattern=north east lines, pattern color=black!75] (1,1) rectangle +(1,1);}
        \modpattern{}{2,1}{0/0,0/2,1/0,2/1}
        {\only<2> \filldraw (1.66,1.33) circle (3pt);}
        {\only<2> \filldraw (1.33,1.66) circle (3pt);}
        {\only<3> \filldraw (2.66,0.33) circle (3pt);}
        {\only<3> \filldraw (2.33,0.66) circle (3pt);}
    \end{tikzpicture}
  \end{equation*}
  {\only<1-3> \vphantom{This is \(m_2\).}}
  {\only<4-> {}}
  \end{example}
\end{frame}

>>>>

\begin{frame}{Second Dominating rule}
  \begin{block}{}
    The patterns
    \begin{equation*}
    m_1 = \pattern{}{2,1}{0/0,0/1,0/2,1/0,2/1} \text{ and } m_2 = \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/1,2/1}
    \end{equation*}
    are coincident in \(\av{\perm{2,3,1}}\).
  \end{block}
\end{frame}

@@

\begin{frame}
  \begin{example}
  Considering \(m_1\) again
  \begin{equation*}
    \begin{tikzpicture}[scale=\picscale]
        {\only<3-4> \fill[pattern=north east lines, pattern color=black!75] (1,1) rectangle +(1,1);}
        \modpattern{}{2,1}{0/0,0/1,0/2,1/0,2/1}
        {\only<2> \filldraw (1.66,1.33) circle (3pt);}
        {\only<2> \filldraw (1.33,1.66) circle (3pt);}
        {\only<3> \filldraw (2.66,0.33) circle (3pt);}
        {\only<3> \filldraw (2.33,0.66) circle (3pt);}
    \end{tikzpicture}
  \end{equation*}
  {\only<1-3> \vphantom{This is \(m_2\).}}
  {\only<4-> {}}
  \end{example}
\end{frame}

### 2016-05-26 writing/presentation/confut.tex
Changes to presentation from Henning's notes
Make changes suggested by Michael to thesis. Bjarki's pedantry in presentation.

\section{Conclusions and Future Work}
\label{sec:Conclusions and Future Work}
\subsection{Conclusions}
\label{sub:Conclusions}
\begin{frame}{Conclusions}
  \begin{itemize}
  \item Automatic coincidence classification is difficult.
  \begin{itemize}
    \item Completely classified for length 2 mesh patterns and length 3 classical patterns.
    \item Gets harder with larger dominating patterns.
  \end{itemize}
  \item There are a number of Wilf-classes that give interesting
  enumerations.
  \begin{itemize}
    \item Would be interesting to try and connect these sets to different objects.
  \end{itemize}
  \end{itemize}
\end{frame}
\subsection{Future Work}
\label{sub:Future Work}
\subsubsection{Extensions of rules}
\label{subs:Extensions of rules}
\begin{frame}{Future Work: Extensions of rules}
    One can consider application of the Third Dominating rule, as well as a
    simple extension, without any dominating pattern. This can capture the special
    case described by Claesson, Tenner, and Ulfarsson in \cite{ABH}
    \begin{equation*}
     m_1 = \pattern{}{1,2}{0/1,0/2,1/1,1/2,2/0}, \text{ and } m_2 = \pattern{}{1,2}{0/2,1/0,1/1,2/0,2/1}
   \end{equation*}

@@

It is also interesting to consider Wilf-equivalence when we have different dominating
patterns
We can show that the sets \(\mathcal{T} = \av{\textpattern{}{1,2}{1/2,2/0,0/0,1/0,1/1,2/1},\perm{2,3,1}}\) and
\(\mathcal{U} = \av{\textpattern{}{1,2}{1/2,2/0,0/0,1/0,1/1,2/1},\perm{3,2,1}}\)
are Wilf-equivalent.
\pause
\begin{equation*}
    \raisebox{0.6ex}{
    \begin{tikzpicture}[scale=1, baseline=(current bounding box.center)]
        \draw (0.5,1.5) node {\(b_1\)};
        \draw (0.5,2.5) node {\( b_2\)};
        \draw (2.5,2.5) node {\( b_3\)};
        \modpattern[2]{2}{1,2}{0/0,1/0,1/1,1/2,2/0,2/1}
    \end{tikzpicture}
    }
\end{equation*}
\end{frame}

>>>>

\section{Conclusions and Future Work}
\label{sec:Conclusions and Future Work}
\subsection{Conclusions}
\label{sub:Conclusions}
\begin{frame}{Conclusions}
  \begin{itemize}
  \item Automatic coincidence classification is difficult.
  \begin{itemize}
    \item Completely classified for length 2 mesh patterns and length 3 classical patterns.
    \item Gets harder with larger dominating patterns.
  \end{itemize}
  \item There are a number of Wilf-classes that give interesting
  enumerations.
  \begin{itemize}
    \item Would be interesting to try and connect these sets to different objects.
  \end{itemize}
  \end{itemize}
\end{frame}
\subsection{Future Work}
\label{sub:Future Work}
\subsubsection{Extensions of rules}
\label{subs:Extensions of rules}
\begin{frame}{Future Work: Extensions of rules}
    One can consider application of the Third Dominating rule, as well as a
    simple extension, without any dominating pattern. This can capture the special
    case described by Claesson, Tenner, and Ulfarsson in \cite{ABH}, that was
    unexplained by the general results proved in that paper.
    \begin{equation*}
     m_1 = \pattern{}{1,2}{0/1,0/2,1/1,1/2,2/0}, \text{ and } m_2 = \pattern{}{1,2}{0/2,1/0,1/1,2/0,2/1}
   \end{equation*}

@@

It is also interesting to consider Wilf-equivalence when we have different dominating
patterns.
We can show that the sets \(\mathcal{T} = \av{\textpattern{}{1,2}{1/2,2/0,0/0,1/0,1/1,2/1},\perm{2,3,1}}\) and
\(\mathcal{U} = \av{\textpattern{}{1,2}{1/2,2/0,0/0,1/0,1/1,2/1},\perm{3,2,1}}\)
are Wilf-equivalent.
\pause
\begin{equation*}
    \raisebox{0.6ex}{
    \begin{tikzpicture}[scale=1, baseline=(current bounding box.center)]
        \draw (0.5,1.5) node {\(b_1\)};
        \draw (0.5,2.5) node {\( b_2\)};
        \draw (2.5,2.5) node {\( b_3\)};
        \modpattern[2]{2}{1,2}{0/0,1/0,1/1,1/2,2/0,2/1}
    \end{tikzpicture}
    }
\end{equation*}
\end{frame}

### 2016-05-26 writing/presentation/wilf.tex
Make changes suggested by Michael to thesis. Bjarki's pedantry in presentation.

\end{frame}
\subsection{Generating Functions}
\label{sub:Generating Functions}
\begin{frame}{Generating {\huge FUN}ctions}
  \begin{block}{}
    \blockquote[Herbert Wilf, \emph{generatingfunctionology}~\cite{Wilf}]{
    A generating function is a clothesline on which we hang up a sequence of numbers for display.}
    % \flushright -Herbert Wilf, \emph{generatingfunctionology}
  \end{block}
  \begin{definition}
    The \emph{ordinary generating function (OGF)} of a sequence \((A_n)\) is the
    formal power series
    \begin{equation}
      A(x) = \sum_{n=0}^{\infty} A_nx^n
    \end{equation}
    The right hand side is the Taylor expansion of the left hand side at \(0\).
  \end{definition}
\end{frame}
\begin{frame}
  \begin{example}
    The generating function for the constant sequence \(\{1,1,1,1,\dotsc\}\) is
    \begin{equation*}
      f(x) = \frac{1}{1-x}
    \end{equation*}
  \end{example}
\end{frame}
\begin{frame}
  \begin{example}
  The generating function for the avoiders of the permutation \(\perm{2,3,1}\)
  satisfies
  \begin{equation}
    C(x) = 1 +xC^2(x)
  \end{equation}
  This can be seen structurally (\(\mathcal{C}\) is the set of \(\perm{2,3,1}\) avoiders)
  \begin{figure}[!ht]
    \centering
    \(\mathcal{C} = \varepsilon \mathrel{\sqcup}\) \decompmax{0/1,1/0}{\only<1>A\only<2->{\mathcal{C}}}{}{}{\only<1>B\only<2->{\mathcal{C}}}
\end{figure}
\only<3->{
Solving for \(C(x)\) and taking the Taylor expansion at \(0\) gives
\begin{equation*}
  1+x+2x^2+5x^3+14x^4+42x^5+132x^6+1430x^7+4862x^8+\dotsb
\end{equation*}
}

@@

\begin{frame}
  First we consider the avoiders \(\mathcal{M}_1\) of \(m_1\) in \(\av{\perm{2,3,1}}\).
  \begin{equation*}
    \mathcal{M}_1 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\mathcal{M}_1}{}{}{\mathcal{M}^\prime_1}
\end{equation*}
Where \(\mathcal{M}^\prime_1\) is a permutation avoiding \(\perm{2,3,1}, m_1\)
and \(\textpattern{}{1}{0/0,0/1,1/0}\).\pause Now considering a permutation in \(\mathcal{M}^\prime_1\)
\begin{equation*}
    \mathcal{M}^\prime_1 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\scriptstyle \mathcal{M}_1\setminus\varepsilon}{}{}{\mathcal{M}^\prime_1}
\end{equation*}
\end{frame}

\begin{frame}
  Next we consider the avoiders \(\mathcal{M}_2\) of \(m_2\) in \(\av{\perm{2,3,1}}\).
  \begin{equation*}
    \mathcal{M}_2 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\mathcal{M}^\prime_2}{}{}{\mathcal{M}_2}
\end{equation*}
Where \(\mathcal{M}^\prime_2\) is a permutation avoiding \(\perm{2,3,1}, m_2\)
and \(\textpattern{}{1}{0/0,0/1,1/1}\).\pause Now considering a permutation in \(\mathcal{M}^\prime_2\)
\begin{equation*}
    \mathcal{M}^\prime_2 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\mathcal{M}^\prime_2}{}{}{\scriptstyle \mathcal{M}_2\setminus\varepsilon}
\end{equation*}
\end{frame}
\begin{frame}
  Both of these sets therefore have generating function \(M(x)\) satisfying
  \begin{align}
    M(x) &= 1 + xM(x)M^\prime(x) \label{eqn:Pgen}\\
    M^\prime(x) &= 1 + x(M(x)-1)M^\prime(x)\label{eqn:PprimeGen}
 \end{align}\pause
 Solving \ref{eqn:PprimeGen} for \(M^\prime(x)\) and substituting into
\ref{eqn:Pgen} gives us that the generating function for
\(M(x)\) satisfies
\begin{equation}
    M(x) = xM^2(x) - x(M(x) - 1) + 1
\end{equation}\pause
Evaluating coefficients of \(M(x)\) then gives the sequence
\begin{equation*}
    1, 1, 1, 2, 4, 9, 21, 51, 127, 323, 835,\dotsc
\end{equation*}
This is an offset of the Motzkin numbers.
\end{frame}

>>>>

\end{frame}
\subsection{Generating Functions}
\label{sub:Generating Functions}
\begin{frame}{Generating {\huge FUN}ctions}
  \begin{block}{}
    \blockquote[Herbert Wilf, \emph{generatingfunctionology}~\cite{Wilf}]{%
    A generating function is a clothesline on which we hang up a sequence of numbers for display.}
    % \flushright -Herbert Wilf, \emph{generatingfunctionology}
  \end{block}
  \begin{definition}
    The \emph{ordinary generating function (OGF)} of a sequence \((A_n)\) is the
    formal power series
    \begin{equation}
      A(x) = \sum_{n=0}^{\infty} A_nx^n
    \end{equation}
    The right hand side is the Taylor expansion of the left hand side at \(0\).
  \end{definition}
\end{frame}
\begin{frame}
  \begin{example}
    The generating function for the constant sequence \(\{1,1,1,1,\dotsc\}\) is
    \begin{equation*}
      f(x) = \frac{1}{1-x}
    \end{equation*}
  \end{example}
\end{frame}
\begin{frame}
  \begin{example}
  The generating function for the avoiders of the permutation \(\perm{2,3,1}\)
  satisfies
  \begin{equation}
    C(x) = 1 +xC^2(x)
  \end{equation}
  This can be seen structurally (\(\mathcal{C}\) is the set of \(\perm{2,3,1}\) avoiders)
  \begin{figure}[!ht]
    \centering
    \(\mathcal{C} = \varepsilon \mathrel{\sqcup}\) \decompmax{0/1,1/0}{\only<1>A\only<2->{\mathcal{C}}}{}{}{\only<1>B\only<2->{\mathcal{C}}}
\end{figure}
\only<3->{
Solving for \(C(x)\) and taking the Taylor expansion at \(0\) gives
\begin{equation*}
  1+x+2x^2+5x^3+14x^4+42x^5+132x^6+1430x^7+4862x^8+\dotsb
\end{equation*}
}

@@

\begin{frame}
  First we consider the avoiders \(\mathcal{M}_1\) of \(m_1 = \pattern{}{1,2}{1/0,1/1,1/2,2/0,2/1} \) in \(\av{\perm{2,3,1}}\).
  \begin{equation*}
    \mathcal{M}_1 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\mathcal{M}_1}{}{}{\mathcal{M}^\prime_1}
\end{equation*}
Where \(\mathcal{M}^\prime_1\) is a permutation avoiding \(\perm{2,3,1}, m_1\)
and \(\textpattern{}{1}{0/0,0/1,1/0}\). \pause Now considering a permutation in \(\mathcal{M}^\prime_1\)
\begin{equation*}
    \mathcal{M}^\prime_1 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\scriptstyle \mathcal{M}_1\setminus\varepsilon}{}{}{\mathcal{M}^\prime_1}
\end{equation*}
\end{frame}

\begin{frame}
  Next we consider the avoiders \(\mathcal{M}_2\) of \(m_2 = \pattern{}{2,1}{0/1,1/0,1/1,1/2,2/1}\) in \(\av{\perm{2,3,1}}\).
  \begin{equation*}
    \mathcal{M}_2 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\mathcal{M}^\prime_2}{}{}{\mathcal{M}_2}
\end{equation*}
Where \(\mathcal{M}^\prime_2\) is a permutation avoiding \(\perm{2,3,1}, m_2\)
and \(\textpattern{}{1}{0/0,0/1,1/1}\). \pause Now considering a permutation in \(\mathcal{M}^\prime_2\)
\begin{equation*}
    \mathcal{M}^\prime_2 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\mathcal{M}^\prime_2}{}{}{\scriptstyle \mathcal{M}_2\setminus\varepsilon}
\end{equation*}
\end{frame}
\begin{frame}
  Both of these sets therefore have generating function \(M(x)\) satisfying
  \begin{align}
    M(x) &= 1 + xM(x)M^\prime(x) \label{eqn:Pgen}\\
    M^\prime(x) &= 1 + x(M(x)-1)M^\prime(x)\label{eqn:PprimeGen}
 \end{align}\pause
 Solving \ref{eqn:PprimeGen} for \(M^\prime(x)\) and substituting into
\ref{eqn:Pgen} gives us that the generating function for
\(M(x)\) satisfies
\begin{equation}
    M(x) = xM^2(x) - x(M(x) - 1) + 1
\end{equation}\pause
Evaluating coefficients of \(M(x)\) then gives the sequence
\begin{equation*}
    1, 1, 1, 2, 4, 9, 21, 51, 127, 323, 835,\dotsc
\end{equation*}
This is an offset of the Motzkin numbers.
\end{frame}
